\large How resource abundance and stochasticity affect animals' spatial needs

### 2023-08-06 writing/manuscript.Rmd
reorganized application section

# Applying the framework: The effects of $\e(R)$ and $\var(R)$ on a lowland tapir

\noindent In this section, we use empirical tracking data on a lowland tapir from the Brazilian Cerrado to show how this framework can be applied to better understand the movement ecology of free-ranging animals via remote sensing data such as NDVI [@pettorelli_normalized_2011]. Appendix 2 contains additional information on how NDVI was modeled along with details on the continuous-time movement models [@fleming_ctmm_2021; @noonan_scale-insensitive_2019] and autocorrelated kernel density estimation [@noonan_comprehensive_2019; @alston_mitigating_2022; @silva_autocorrelationinformed_2022] used to quantify the tapir's spatial needs.

@@

We start by providing examples and considerations on how one may model $R$, animals' spatial needs, and the effects of $\e(R)$ and $\var(R)$ on animals' spatial needs. Next, we apply the methods to the tapir's GPS tracking data and use NDVI as a proxy for $R$. Finally, we offer suggestions on how this approach can be used to inform conservation-related decisions, including assessing habitat quality and estimating animals' spatial needs under different scenarios.

@@

### Estimating resource abundance using NDVI

@@

### Modeling the effects of $\e(R)$ and $\var(R)$

@@

### Empirical results

```{r tapir-mw, fig.cap="Seven-day home-range size of a lowland tapir (\\emph{Tapirus terrestris}) in response to changes in mean and variance in resource abundance. (a.) Mean resource abundance, $\\mu(t)$, esimated as the mean NDVI at the locations visited by the tapir. (b.) Varince in resource abundance, $\\sigma^2(t)$, esimated as the average variance in NDVI at the locations visited by the tapir. (c.) Estimated seven-day home range based on the 95\\% utilization quantiles. (d., e.) estimated marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ on home-range size. The model accounted for the marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ and their interaction on mean spatial needs and the variance around them. (f.) The effect of $\\mu(t)$ does not follow the data closely because $\\e(R)$ and $\\var(R)$ are highly correlated. Consequently, while estimating the effects of $\\e(R)$ and $\\var(R)$ via separate models would allow result in a closer fit, the estimated effects would be inappropriate because they do not disentangle the effects of $\\e(R)$ and $\\var(R)$. See Appendix 2 for additional information. The tapir movement data corresponds to the individual named ``Anna\" from the Cerrado sample of Medici $et~al.$ (2022).", out.width='100%'}
knitr::include_graphics('../figures/tapir-example.png')
```

>>>>

# Applying the framework

\noindent In this section, we show how this framework can be applied to better understand the movement ecology of free-ranging animals via remote sensing data such as NDVI [@pettorelli_normalized_2011]. To illustrate the methods, we use empirical tracking data on a lowland tapir from the Brazilian Cerrado. Appendix 2 contains additional information on how NDVI was modeled along with details on the continuous-time movement models [@fleming_ctmm_2021; @noonan_scale-insensitive_2019] and autocorrelated kernel density estimation [@noonan_comprehensive_2019; @alston_mitigating_2022; @silva_autocorrelationinformed_2022] used to quantify the tapir's spatial needs.

@@

We start by providing examples and considerations on how one may model $R$ and the effects of $\e(R)$ and $\var(R)$ on animals' spatial needs. Next, we apply the methods to the tapir's GPS tracking data and use NDVI as a proxy for $R$. Finally, we offer suggestions on how this approach can be used to inform conservation-related decisions, including assessing habitat quality and estimating animals' spatial needs under different scenarios.

@@

### Estimating $R$ using NDVI

@@

### Modeling the effects of $\e(R)$ and $\var(R)$ on animals' spatial needs

@@

### The effects of $\e(R)$ and $\var(R)$ on a lowland tapir's spatial needs

```{r tapir-mw, fig.cap="Seven-day home-range size of a lowland tapir (\\emph{Tapirus terrestris}) in response to changes in mean and variance in resource abundance. (a.) Mean resource abundance, $\\mu(t)$, esimated as the mean NDVI at the locations visited by the tapir. (b.) Varince in resource abundance, $\\sigma^2(t)$, esimated as the average variance in NDVI at the locations visited by the tapir. (c.) Estimated seven-day home range based on the 95\\% utilization quantiles. (d., e.) estimated marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ on home-range size. The model accounted for the marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ and their interaction on mean spatial needs and the variance around them. (f.) The effect of $\\mu(t)$ does not follow the data closely because $\\e(R)$ and $\\var(R)$ are highly correlated. Consequently, while estimating the effects of $\\e(R)$ and $\\var(R)$ via separate models would allow result in a closer fit, the estimated effects would be inappropriate because they do not disentangle the effects of $\\e(R)$ and $\\var(R)$. See Appendix 2 for additional information. The tapir movement data corresponds to the individual named ``Anna\" from the Cerrado sample of Medici $et~al.$ (2022).", out.width='100%'}
knitr::include_graphics('../figures/tapir-example-with-data.png')
```
